```{r}
install.packages("Hmisc")
install.packages("segmented")

library(segmented)
library(tidyverse)  
library(Hmisc)
```


```{r prepare_data}
bat_data <- read_csv("dataset/MLB_bat_data_2021.csv") %>% 
  group_by(player_id) %>% 
  arrange(game_date) %>% 
  mutate(temp =1, games = cumsum(temp), plate_appearances = sum(pa)) %>%
  mutate(hr_total=sum(home_runs) , hit_total=sum(h) , bb_total=sum(bb) , so_total=sum(so)) %>% 
  mutate(hr_cum=cumsum(home_runs), hit_cum=cumsum(h), bb_cum=cumsum(bb), so_cum=cumsum(so)) %>% 
  filter(plate_appearances>502)

## filter for qualified players = 502 plate apperances
```


```{r calculate_paces_function}
segment_model <- function(batter, stat, dataset){
  if (dataset == "MLB")      {d1 <- bat_data %>% filter(player==batter)} 
  else if (dtatset == "NBA") {d1 <- offense_data %>% filter(player==batter)} 
  else                       {print("invalid dataset")}
  
  if (stat == "hr")          {fit_lm <- lm(hr_cum ~ 1 + games, data = d1)
                             total <- max(d1$hr_total)}
  else if (stat =="hit")    {fit_lm <- lm(hit_cum ~ 1 + games, data = d1)
                             total <- max(d1$hit_total)}
  else if (stat =="bb")      {fit_lm <- lm(bb_cum ~ 1 + games, data = d1)
                             total <- max(d1$bb_total)}
  else if (stat =="so")      {fit_lm <- lm(so_cum ~ 1 + games, data = d1)
                             total <- max(d1$so_total)}
  else                       {print("invalid stat")}
  
  #fit_lm <- lm(hr_cum ~ 1 + games, data = d1)
  fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
  
  #total <- max(d1$hr_total)
  total_games <- max(d1$games)
  slope <- total/total_games
  
  team = max(d1$team_abbrev)
  
  a <- fit_segmented$psi
  n <- length(a)/3
  breakpoint <- c(0, a[(n+1):(2*n)], max(d1$games))
  b <- fit_segmented$coefficients[2:(n+2)]
  pace <- cumsum(b)
  pace_length <- (breakpoint - lag(breakpoint))[2:(n+2)]
  weighted_pace <- weighted.mean(pace, pace_length)
  tibble(name = batter, team = team, total = total, total_games = total_games, slope  = slope, weighted_pace = weighted_pace, pace = pace, pace_length = pace_length)
}
```

```{r caclulate_hr_paces}
## HOMERUNS
  hr_data <- bat_data %>% filter(hr_total>15)
  hr_paces <- map_dfr(unique(hr_data$player), segment_model, stat = "hr", dataset = "MLB")
  
  hr_streakiness <- hr_paces %>% 
    group_by(name) %>% 
    summarise(total=mean(total), streak=sqrt(wtd.var(pace,pace_length)))

  exp <- lm(total ~ streak, data=hr_streakiness)
  summary(exp)
  beta <- coef(exp)

  ggplot(hr_streakiness) +
    geom_point(aes(x=streak, y=total)) +
    geom_abline(intercept=beta[1], slope=beta[2], color="red")
```

```{r caclulate_hit_paces}
## HITS
  hit_paces <- map_dfr(unique(bat_data$player), segment_model, stat = "hit", dataset = "MLB")
  
  hit_streakiness <- hit_paces %>% 
    group_by(name) %>% 
    summarise(total=mean(total), streak=sqrt(wtd.var(pace,pace_length)))

  exp <- lm(total ~ streak, data=hit_streakiness)
  summary(exp)
  beta <- coef(exp)

  ggplot(hit_streakiness) +
    geom_point(aes(x=streak, y=total)) +
    geom_abline(intercept=beta[1], slope=beta[2], color="red")
```

```{r caclulate_so_paces}
## STRIKE OUTS
  so_paces <- map_dfr(unique(bat_data$player), segment_model, stat = "so", dataset = "MLB")
  
  so_streakiness <- so_paces %>% 
    group_by(name) %>% 
    summarise(total=mean(total), streak=sqrt(wtd.var(pace,pace_length)))

  exp <- lm(total ~ streak, data=so_streakiness)
  summary(exp)
  beta <- coef(exp)

  ggplot(so_streakiness) +
    geom_point(aes(x=streak, y=total)) +
    geom_abline(intercept=beta[1], slope=beta[2], color="red")
```

```{r caclulate_bb_paces}
## WALKS
  bb_paces <- map_dfr(unique(bat_data$player), segment_model, stat = "bb", dataset = "MLB")
  
  bb_streakiness <- bb_paces %>% 
    group_by(name) %>% 
    summarise(total=mean(total), streak=sqrt(wtd.var(pace,pace_length)))

  exp <- lm(total ~ streak, data=bb_streakiness)
  summary(exp)
  beta <- coef(exp)

  ggplot(bb_streakiness) +
    geom_point(aes(x=streak, y=total)) +
    geom_abline(intercept=beta[1], slope=beta[2], color="red")
```
  
```{r player_dataframe}
player_dataframe <- function(batter, stat, dataset){
  if (dataset == "MLB")      {d1 <- bat_data %>% filter(player==batter)} 
  else if (dtatset == "NBA") {d1 <- offense_data %>% filter(player==batter)} 
  else                       {print("invalid dataset")}
  
  if (stat == "hr")          {fit_lm <- lm(hr_cum ~ 1 + games, data = d1)
                             total <- max(d1$hr_total)}
  else if (stat =="hit")    {fit_lm <- lm(hit_cum ~ 1 + games, data = d1)
                             total <- max(d1$hit_total)}
  else if (stat =="bb")      {fit_lm <- lm(bb_cum ~ 1 + games, data = d1)
                             total <- max(d1$bb_total)}
  else if (stat =="so")      {fit_lm <- lm(so_cum ~ 1 + games, data = d1)
                             total <- max(d1$so_total)}
  else                       {print("invalid stat")}
  
  fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
  
  team = max(d1$team_abbrev)
  
  my_fitted <- fitted(fit_segmented)
  data.frame(player = batter, team = team, games = d1$games, stat_cum = my_fitted)
}
```

```{r individual_dataframes}
  bb_datframe <- map_dfr(unique(bat_data$player), player_dataframe, stat = "bb", dataset = "MLB")

  so_datframe <- map_dfr(unique(bat_data$player), player_dataframe, stat = "so", dataset = "MLB")
  
  hit_datframe <- map_dfr(unique(bat_data$player), player_dataframe, stat = "hit", dataset = "MLB")
  
  hr_datframe <- map_dfr(unique(hr_data$player), player_dataframe, stat = "hr", dataset = "MLB")
```

  
```{r view_individual}
d1 <- bat_data %>% filter(player_id=="perezsa02")
# d1 <- bat_data %>% filter(player_id=="crawfbr01")
# d1 <- bat_data %>% filter(player_id== "hanigmi01")

# Home run fit
fit_lm <- lm(hr_cum ~ 1 + games, data = d1)
fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
my_fitted <- fitted(fit_segmented)
hr_model <- data.frame(games = d1$games, hr_cum = my_fitted)

# hit fit
fit_lm <- lm(hit_cum ~ 1 + games, data = d1)
fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
my_fitted <- fitted(fit_segmented)
hit_model <- data.frame(games = d1$games, hit_cum = my_fitted)

# strike outs model
fit_lm <- lm(so_cum ~ 1 + games, data = d1)
fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
my_fitted <- fitted(fit_segmented)
so_model <- data.frame(games = d1$games, so_cum = my_fitted)

# walks model
fit_lm <- lm(bb_cum ~ 1 + games, data = d1)
fit_segmented = segmented(fit_lm, seg.Z = ~games, psi=NA, control=seg.control(fix.npsi=FALSE, n.boot=0, tol=1e-7, it.max = 50))
my_fitted <- fitted(fit_segmented)
bb_model <- data.frame(games = d1$games, bb_cum = my_fitted)

p1 <- ggplot() + 
  geom_line(data = hr_model, aes(x = games, y = hr_cum)) + 
  geom_point(data = d1, aes(x=games, y=hr_cum), color="yellow", alpha =0.1)

p2 <- ggplot() +
  geom_line(data = hit_model, aes(x = games, y = hit_cum)) + 
  geom_point(data = d1, aes(x=games, y=hit_cum), color="blue", alpha =0.1)

p3 <- ggplot() +
  geom_line(data = so_model, aes(x = games, y = so_cum)) + 
  geom_point(data = d1, aes(x=games, y=so_cum), color="red", alpha =0.1)

p4 <- ggplot() +
  geom_line(data = bb_model, aes(x = games, y = bb_cum)) + 
  geom_point(data = d1, aes(x=games, y=bb_cum), color="green", alpha =0.1)

grid.arrange(p1, p2, p3, p4, ncol=2)