---
title: "Leo_EDA_2"
author: "Leo Seoane"
date: "2022-11-09"
output: html_document
---


##what teams have the most players with streaks

##Are the trends meaningfully different if we look at points, rebounds, and assists? 

##How are players likely to perform relative to their last 5 performances? 
##Visualize by how much players perform above average in games that they do well

##analyze how players perform against teams much better than them

##pick top 5, middle 5, bottom 5 players to display

##for visualization, show how changing ts pct,usg pct, off rtg changes probability of changing prediction maybe using sliding value


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(modelr))
```
```{r data loading}
load(here::here("dataset-ignore", "clean_nba_data.RData"))
data <-data %>% mutate(PRA=pts+trb+ast) %>%mutate(usg_pct=usg_pct/100)
problems()

```

```{r Final player EDA}
#get all of our data by players who meet criteria and calculate streaks
all_data <-data %>% group_by(player)  %>% filter(mean(minutes) >25,minutes >15) %>% mutate(change_net = pts - mean(pts),pts_direction  = 
sign(change_net),change_net_PRA=PRA - mean(PRA),PRA_direction=sign(change_net_PRA)) %>%mutate(pts_streak=ave(pts_direction, cumsum(pts_direction==-1), FUN = seq_along) - 1,PRA_streak=ave(PRA_direction, cumsum(PRA_direction==-1), FUN = seq_along) - 1)

#visualize distribution of positive player points streaks and PRA streaks
streak <-all_data$pts_streak
indices<-which(diff(sign(diff(streak)))==-2)+1
local_max <-slice(all_data,indices)
local_max<-local_max %>%group_by(player) %>% select(pts_streak,change_net,change_net_PRA,PRA_streak) %>%filter(pts_streak >0,PRA_streak >0)
summary(local_max)
ggplot(local_max) +geom_freqpoly(aes(x=pts_streak))
ggplot(local_max) +geom_freqpoly(aes(x=PRA_streak))


#view proportion of continuing streaks
best_streaks_2022 <- all_data %>% summarize(max_streak=max(pts_streak)) %>% arrange(desc(max_streak)) %>%head(10)
top_player_data22 <- all_data %>% filter(player %in% best_streaks_2022$player) %>%mutate(pts_direction=ifelse(pts_direction==-1,0,1))
top_player_data22 <-top_player_data22 %>% group_by(player) %>% mutate(a=10*lag(pts_direction)+pts_direction) %>%ungroup() 
top_player_proportions <-top_player_data22 %>% mutate(a=str_c(lag(pts_direction),pts_direction,sep = ',')) %>% count(a) %>% group_by(a)
top_player_proportions

#cdf and pdf for player points
ggplot(best_streaks_2022,aes(max_streak))+stat_ecdf(geom = "step")
ggplot(best_streaks_2022,aes(max_streak))+geom_density()
ggplot(top_player_data22,aes(pts)) +geom_density()

#model building
mod1 <-all_data %>%ungroup() %>% glm(formula=factor(pts_direction)~ts_pct +off_rtg+usg_pct+usg_pct:ts_pct,family=binomial())
mod1$coefficients
exp(mod1$coefficients)

all_data <-all_data %>%mutate(pts_direction=ifelse(pts_direction==-1,0,1))
grid <- data_grid(ungroup(all_data), ts_pct= seq_range(ts_pct,100),off_rtg=seq_range(off_rtg,3),usg_pct= seq_range(usg_pct,3)) %>% add_predictions(mod1, type = "response" )

ggplot(all_data) +geom_line(aes(x=ts_pct,y=pred, color = factor(usg_pct), linetype = factor(off_rtg)),data=grid)

```