library(tidyverse)
library(here)
install.packages('broom')
##what teams have the most players with streaks

##Are the trends meaningfully different if we look at points, rebounds, and assists? 

##How are players likely to perform relative to their last 5 performances? 
##Visualize by how much players perform above average in games that they do well

##analyze how players perform against teams much better than them

```{r data loading}
data <-read_csv(here::here("dataset-ignore/raw_nba_data.csv"),show_col_types = FALSE)
#problems()
```


```{r question 1}
##what is the average length of a streak in the nba by player and team?

#by player
all_data <-data %>% group_by(player) %>% filter(mean(minutes) >25,minutes >15) %>% mutate(change_net = pts - mean(pts),direction  = 
sign(change_net)) %>% mutate(streak = cumsum(direction = lag(direction, default = 0))) 


streak <-all_data$streak
indices<-which(diff(sign(diff(streak)))==-2)+1
local_max <-slice(all_data,indices)
local_max_all <- local_max %>%group_by(player) %>% select(streak)
local_max_positive <-local_max %>%group_by(player) %>% select(streak,change_net) %>%filter(streak >-1)

summary(local_max_all$streak)
summary(local_max_positive$streak)



#by team
team_data <-data  %>% distinct(game_id,.keep_all = TRUE) %>%group_by(Team_Abbrev) %>% mutate(change_net = Team_Score - mean(Team_Score),direction  = 
                                                                                              sign(change_net)) %>% mutate(streak = cumsum(direction = lag(direction, default = 0))) 

team_streak <-team_data$streak
indices<-which(diff(sign(diff(team_streak)))==-2)+1
team_local_max <-slice(team_data,indices)
team_local_max_all <- team_local_max %>%group_by(player) %>% select(streak)
team_local_max_positive <-team_local_max %>%group_by(player) %>% select(streak,change_net) %>%filter(streak >-1)

summary(team_local_max_all$streak)
summary(team_local_max_positive$streak)

#distribution of streaks by player 
ggplot(local_max_all) +geom_freqpoly(aes(x=streak))
ggplot(local_max_positive) +geom_freqpoly(aes(x=streak))

#distribution of streaks by team
ggplot(team_local_max_all) +geom_freqpoly(aes(x=streak))
ggplot(team_local_max_positive) +geom_freqpoly(aes(x=streak))



```

```{r question 2}

##what are the highest streaks in the nba by player 2022
best_streaks_2022 <- data %>% group_by(player) %>% filter(mean(minutes) >25,season==2022,minutes >15) %>% mutate(change_net = pts - mean(pts),direction  = sign(change_net))  %>%mutate(streak = cumsum(direction = lag(direction, default = 0))) %>% summarize(max_streak=max(streak)) %>% arrange(desc(max_streak)) %>%head(10)


top_player_data22 <- data %>% filter(player %in% best_streaks_2022$player) %>% group_by(player)%>% mutate(change_net = pts - mean(pts),
direction  = sign(change_net)) %>%mutate(usg_pct=usg_pct/100) %>%mutate(direction=ifelse(direction==-1,0,1))

best_player_streaks_2022 <- data %>% group_by(player) %>% filter(mean(minutes) >25,season==2022,minutes >15) %>% mutate(change_net = pts - mean(pts),direction  = sign(change_net))  %>%mutate(streak = cumsum(direction = lag(direction, default = 0))) 

##chance that top player streaks continue
##What are the likelihoods that a player is going to continue a streak once they have started
a_player_data_agg <-top_player_data22 %>% mutate(a=10*lag(direction)+direction) 
a_player_data_agg <-a_player_data_agg %>% mutate(a=str_c(lag(direction),direction,sep = ',')) %>% count(a) %>% mutate(prop=n/sum(n))

a_player_data <-best_player_streaks_2022 %>% mutate(a=10*lag(direction)+direction)
a_player_data <-a_player_data%>% mutate(a=str_c(lag(direction),direction,sep = ',')) %>% count(a) %>% mutate(prop=n/sum(n))


knitr::kable(a_player_data_agg)
knitr::kable(a_player_data)

#by team
best_team_streaks_2022 <- data %>% distinct(game_id,.keep_all=TRUE) %>% group_by(Team_Abbrev) %>%  mutate(change_net = Team_Score - mean(Team_Score),direction  = sign(change_net))  %>%mutate(streak = cumsum(direction = lag(direction, default = 0))) %>% mutate(max_streak=max(streak))

top_team_data22 <- data %>% filter(Team_Abbrev %in% best_team_streaks_2022$Team_Abbrev) %>% mutate(change_net = Team_Score - mean(Team_Score),
direction  = sign(change_net)) %>%mutate(usg_pct=usg_pct/100) %>%mutate(direction=ifelse(direction==-1,1,0))

##chance that team streaks continue
a_team_data_agg <-top_team_data22 %>% ungroup() %>% mutate(a=10*lag(direction)+direction) %>% count(a) 
a_team_data_agg <-top_team_data22 %>% mutate(a=str_c(lag(direction),direction,sep = ',')) %>% count(a) %>% mutate(prop=n/sum(n))

a_team_data <- best_team_streaks_2022 %>% ungroup() %>% mutate(a=10*lag(direction)+direction) %>% count(a) 
a_team_data <- best_team_streaks_2022 %>% mutate(a=str_c(lag(direction),direction,sep = ',')) %>% count(a) %>% mutate(prop=n/sum(n))


knitr::kable(a_team_data)
knitr::kable(a_team_data_agg)

```


```{r model building}
##prediction model using lgr and relevant factors for predicting top player performance
mod1 <-top_player_data22 %>%ungroup() %>% glm(formula=factor(direction)~ts_pct +off_rtg+usg_pct,family=binomial())
mod1 %>%coef()
tableas<-broom::glance(mod1)
grid <- top_player_data22   %>%ungroup() %>%
  mutate(pred = predict(mod1, newdata = ., type = 'response'))


ggplot(grid) +geom_point(aes(y=pred,x=pts,color=direction))+stat_smooth(aes(x=pts,y=pred),method="glm", se=FALSE, fullrange=TRUE, method.args = list(family=binomial))


##prediction model using modified expected equations from popular sports model maker Ken Pomeroy

expected_pos <- function(t1_p,t2_p,la_p){
  t1_p <-t1_p / la_p
  t2_p <-t2_p / la_p
  e_pos <-t1_p *t2_p *la_p
  return(e_pos)
}
expected_output <-function(t1_or,t2_dr,la_or,t1_p,t2_p,la_p){
  t1_out <-(t1_or /100) *(t2_dr /100) *la_or
  e_pos <-expected_pos(t1_p,t2_p,la_p)
  t1_out <-t1_out *(e_pos/100)
  
  return(t1_out)
}

model_data <- data  %>% distinct(game_id,.keep_all = TRUE) %>% group_by(Team_Abbrev) %>% mutate(Team_def_rtg=mean(Opponent_Score/Opponent_pace *100), Opponent_def_rtg=mean(Team_Score/Team_pace *100)) %>%ungroup() %>%mutate(la_pace=mean(Team_pace),la_or=mean(Team_off_rtg))

model_data <-model_data %>% select(Team_off_rtg,Opponent_def_rtg,la_or,Team_pace,Opponent_pace,la_pace)
preds <- expected_output(model_data$Team_off_rtg,model_data$Opponent_def_rtg,model_data$la_or,model_data$Team_pace,model_data$Opponent_pace,model_data$la_pace)
model_data <-model_data %>% cbind(preds)


```
```{r other stat graphs}
#cdf and pdf for player points
ggplot(best_streaks_2022,aes(max_streak))+stat_ecdf(geom = "step")
ggplot(best_streaks_2022,aes(max_streak))+geom_density()
ggplot(top_player_data22,aes(pts)) +geom_density()

#cdf and pdf for team points

ggplot(best_team_streaks_2022,aes(max_streak))+stat_ecdf(geom = "step")
ggplot(best_team_streaks_2022,aes(max_streak))+geom_density()
ggplot(top_team_data22,aes(Team_Score)) +geom_density()
```